A novel method for detecting coalescing binaries in near realtime with Advanced LIGO and beyond \editorial{Submit to PRD or CQG?}

### 2011-02-21 appendix.tex
Fixing latency of FFT method in appendix
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@659 ccdd05d5-56c1-461a-bf98-85f6876523b4

The \textsc{fir} filter implementation has the advantage that it has no intrinsic latency, whereas the \textsc{fft} convolution has at least the latency of the \textsc{fft} block size $D \geq 2 n$.  \editorial{Drew: Should the latency be D-n?} For example, for a $1.4 - 1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the \textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.  However, the \textsc{fir} filter implementation has the disadvantage of much greater overhead per sample than the \textsc{fft} convolution.  For a $1\,\mathrm{ks}$ template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation requires about about $n / 8 \lg 2 n = 2.2 \times 10^4$ times more operations per sample than the \textsc{fft} implementation.

>>>>

The \textsc{fir} filter implementation has the advantage that it has no
intrinsic latency, whereas the \textsc{fft} convolution has at least the
latency of the \textsc{fft} block size $D-n \geq n$. For example, for a $1.4 -
1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the
\textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.  However, the
\textsc{fir} filter implementation has the disadvantage of much greater
overhead per sample than the \textsc{fft} convolution.  For a $1\,\mathrm{ks}$
template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation
requires about about $n / 8 \lg 2 n = 2.2 \times 10^4$ times more operations
per sample than the \textsc{fft} implementation.

### 2011-02-21 method.tex
Fixing latency of FFT method in appendix
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@659 ccdd05d5-56c1-461a-bf98-85f6876523b4

\subsection{Data Whitening}
Matched filtering the SVD basis vectors has motivated a decoupling of the
whitening routine from the matched filtering engine. This was necessary in
order weight templates appropriately by whitening them before the SVD
calculation, and, as such, we have also moved the data whitening outside the
match filter engine. We have chosen to whiten the data using a running
geometric average of the PSD computed by Hann windowing 8 second buffers of
data and using 50\% overlapping buffers. The running geometric average is
updated once per buffer from a median of the recent PSDs. We find that this
algorithm is extremely fast to converge to an accurate PSD and also very robust
against glitches. This is demonstrated below where we have injected
band-limited white noise burst with a central frequency of 150 Hz, a bandwidth
of 100 Hz, a time duration of 0.2 seconds, and a injection density of one
injection every $100/\pi \approx 32$ seconds. At this density we find there
should only ever be one injection affecting the median history so the glitches
should minimally affect PSD estimation as can be seen by comparing the local
RMS amplitude between injections to that without injections.

\begin{center}
\resizebox{0.9\textwidth}{!}{\includegraphics{figures/whiten_burst_step_100_semilogy.png}}
\end{center}

An inspiral, to leading order, has gravitational-wave frequency evolution given
by
\begin{equation}
f(t) = \frac{1}{8 \pi \mathcal{M}} \left( \frac{t_c - t}{5 \mathcal{M}} \right)^{-3/8}\;,
\end{equation}
whose time derivative is given by
\begin{equation}
\frac{df}{dt} = \frac{3}{320 \pi \mathcal{M}^2} \left( \frac{t_c - t}{5 \mathcal{M}} \right)^{-11/8}\;.
\end{equation}
Combining these, we find the frequency evolution as a function of frequency to
be
\begin{equation}
\frac{df}{dt}(f_0) = \frac{3}{320 \pi \mathcal{M}^2} \left( 8 \pi \mathcal{M} f_0 \right)^{3/11}\;,
\end{equation}
which can be inverted to get the minimum frequency which has a given frequency
derivative,
\begin{equation}
f_0 = \frac{1}{8 \pi \mathcal{M}} \left( \frac{320 \pi \mathcal{M}^2}{3} \frac{df}{dt} \right)^{11/3}\;.
\end{equation}

>>>>

\subsection{Data Whitening}
Matched filtering the SVD basis vectors has motivated a decoupling of the
whitening routine from the matched filtering engine. This was necessary in
order weight templates appropriately by whitening them before the SVD
calculation, and, as such, we have also moved the data whitening outside the
match filter engine. We have chosen to whiten the data using a running
geometric average of the PSD computed by Hann windowing 8 second buffers of
data and using 50\% overlapping buffers. The running geometric average is
updated once per buffer from a median of the recent PSDs. We find that this
algorithm is extremely fast to converge to an accurate PSD and also very robust
against glitches.

An inspiral, to leading order, has gravitational-wave frequency evolution given
by
%
\begin{equation}
%
f(t) = \frac{1}{8 \pi \mathcal{M}} \left( \frac{t_c - t}{5 \mathcal{M}}
\right)^{-3/8}\;,
%
\end{equation}
%
whose time derivative is given by
%
\begin{equation}
%
\frac{df}{dt} = \frac{3}{320 \pi \mathcal{M}^2} \left( \frac{t_c - t}{5
\mathcal{M}} \right)^{-11/8}\;.
%
\end{equation}
%
Combining these, we find the frequency evolution as a function of frequency to
be
%
\begin{equation}
%
\frac{df}{dt}(f_0) = \frac{3}{320 \pi \mathcal{M}^2} \left( 8 \pi \mathcal{M}
f_0 \right)^{3/11}\;,
%
\end{equation}
%
which can be inverted to get the minimum frequency which has a given frequency
derivative,
%
\begin{equation}
%
f_0 = \frac{1}{8 \pi \mathcal{M}} \left( \frac{320 \pi \mathcal{M}^2}{3}
\frac{df}{dt} \right)^{11/3}\;.
%
\end{equation}
